---
title: "Column types and coercion"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column types and coercion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


SeaTable bases have `r nrow(harbouR::hb_column_types())` column types. harbouR maps every one
of them onto a predictable R type so the tibble you get back is
analysis-ready, not raw JSON. This vignette is the reference.

## The mapping table

The table below is derived from `hb_column_types()` so it cannot drift
out of sync with the code:

```{r mapping}
library(harbouR)
knitr::kable(hb_column_types())
```

## What the rows look like

`hb_example_rows()` is a small hand-written frame covering the common
types. The bundled `.dtable` fixture is the exhaustive one - it carries a
column of nearly every type - and the last section reads it.

```{r samples}
samples <- hb_example_rows("Samples")
samples
```

### Scalars

```{r scalars}
str(samples[, c("Name", "Concentration", "Status", "Collected")])
```

* `text` / `single-select` come back as `character`.
* `number` becomes a `double` (caveat: SeaTable's numeric column is
  64-bit; large integers may round - if precision matters store them
  as text).
* `date` always becomes `POSIXct`, in UTC. A date-only column simply has
  a zero time component. Call `as.Date()` if you want a calendar date.

### List-columns

These types always come back as list-columns, even when every row has
one value or none. That is the type-stability rule: the same column has
the same R type for every row.

```{r list-types}
subset(hb_column_types(), is_list)$seatable
```

```{r tags}
samples$Tags
```

Flatten with `tidyr::unnest` when you need a long form:

```{r unnest, eval = requireNamespace("tidyr", quietly = TRUE)}
tidyr::unnest(samples[, c("Name", "Tags")], cols = "Tags")
```

### File and image columns

```{r files}
samples$Reports[[1]]
```

Each entry is a list of file objects with `name`, `size`, `type` and
`url` - ready to be downloaded via `hb_download_file()`.

### Read-only columns

These types are maintained by SeaTable, so harbouR drops them from any
write. You can round-trip a tibble through harbouR without writing back
computed values by accident.

```{r read-only}
subset(hb_column_types(), read_only)$seatable
```

Note that `link`, `button` and `digital-sign` are in that list too:
links are maintained through SeaTable's link endpoints rather than by
writing the cell.

## Empty tables

A table with no rows still comes back as a tibble with every column in
its correct type:

```{r empty}
empty <- hb_example_rows("Samples")[0, ]
empty

str(empty, max.level = 1)
```

Every column keeps the type it would have had with rows present -
`Concentration` is still a double, `Collected` still a `POSIXct`, and
`Tags` still a list-column - so downstream code that expects a schema
does not have to special-case the empty result.

## Every type at once

`hb_example_rows()` is deliberately small. The bundled `.dtable` fixture
carries a column of nearly every type SeaTable defines, so it is the
better place to see the whole mapping at work:

```{r fixture}
base <- hb_read_dtable(
  system.file("extdata", "example.dtable", package = "harbouR")
)
schema <- hb_list_columns(base, "Samples")
knitr::kable(schema[, c("name", "type", "editable")])
```

```{r fixture-types}
data <- hb_read_table(base, "Samples")
vapply(data, function(col) class(col)[[1L]], character(1))
```

## Next steps

* `vignette("harbouR")` - the five-minute end-to-end story.
* `vignette("explorer-app")` - exploring a base interactively.
