---
title: "The harbouR explorer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The harbouR explorer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


`hb_run_explorer()` launches a Shiny app for inspecting a SeaTable base
interactively. Reach for it when you want to *see* what is in a base
quickly; reach for the programmatic API (`hb_read_table()`,
`hb_query()` and friends) when you want to *do* something with the
data.

The app and its UI dependencies (`shiny`, `bslib`, `DT`, `reactable`,
`ggplot2`) live in `Suggests`, so the core client stays headless. If
something is missing, the launcher tells you exactly what to install.

## Demo mode

You can try the app right now with no SeaTable account:

```{r demo, eval = FALSE}
harbouR::hb_run_explorer()
```

Then click **Try demo** on the connect screen. The app loads the same
example metadata and rows you saw in the `column-types` vignette -
including list-columns and a `file` column - so every panel is
exercised without a network call.

## The panels

* **Connect.** Server URL and API token, defaulting to `SEATABLE_SERVER`
  and `SEATABLE_API_TOKEN`. **Try demo** bypasses this entirely.
* **Overview.** Value boxes for base name, table count and cached row
  count; a table of the tables, and a bar chart of rows per table.
* **Tables.** Pick a table; the app fetches it via `hb_read_table()`,
  caches it, and shows the typed tibble. A second table lists every
  column's SeaTable type.
* **Query.** A SeaTable SQL console. Results are shown as a table.
  Disabled in demo mode, which has no server to query.

## Launching with a pre-built client

If you already have a `harbour_client`, hand it in and the app starts
connected:

```{r preset, eval = FALSE}
client <- harbouR::hb_client()
harbouR::hb_run_explorer(client)
```

## A note on screenshots

Screenshots for the panels are not bundled with this release - taking
them requires running the app, which `R CMD check` deliberately
prevents. The interactive demo above is the canonical reference.

## Next steps

* `vignette("harbouR")` - the five-minute end-to-end story.
* `vignette("column-types")` - the coercion layer in detail.
