---
title: "Get started with harbouR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with harbouR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

You have a SeaTable base of patient samples and you want to pull it into
R for analysis, push your results back, and attach the resulting PDF
reports - all without leaving R. `harbouR` is an unofficial R client
that lets you do exactly that.

> SeaTable is a trademark of SeaTable GmbH; harbouR is not affiliated
> with or endorsed by SeaTable GmbH.

## Connect

```{r connect, eval = FALSE}
library(harbouR)

client <- hb_client(
  server    = Sys.getenv("SEATABLE_SERVER"),
  api_token = Sys.getenv("SEATABLE_API_TOKEN")
)
```

SeaTable uses a three-token model. You hand harbouR either an
API-token (long-lived, per-base, generated in the SeaTable UI) or a
username+password pair. harbouR exchanges these for a short-lived
**base-token** the first time it needs to, caches it, and refreshes
transparently when it expires - you never see it.

## Inspect the base

The remaining chunks here all run **offline**, against the bundled
example data, so you can knit the vignette without a server.

```{r inspect}
library(harbouR)

meta <- hb_example_metadata()
meta

tibble::as_tibble(meta)
```

## Read a table

Against a live server this is a single call:

```{r read-live, eval = FALSE}
samples <- hb_read_table(client, "Samples")
```

The offline equivalent returns a tibble of the same shape, including
list-columns for multi-select / collaborator / file values:

```{r read-offline}
samples <- hb_example_rows("Samples")
samples
```

## Analyse

`samples` is a regular tibble. Everything you already know about
dplyr / ggplot2 / etc. just works:

```{r analyse, message = FALSE, warning = FALSE}
samples[samples$Status == "ready", c("Name", "Concentration", "Collected")]
```

## Write back

Appending and updating accept any tibble whose columns match the schema:

```{r write-live, eval = FALSE}
new_rows <- tibble::tibble(
  Name = "S-004",
  Concentration = 17.2,
  Status = "draft"
)
hb_append_rows(client, "Samples", new_rows)

hb_update_rows(
  client, "Samples",
  tibble::tibble(`_id` = "r0001", Concentration = 99.9)
)
```

## Attach a file

```{r attach, eval = FALSE}
hb_attach_file(client, "Samples", "r0001", "Reports", "report.pdf")
```

## Next steps

* The `column-types` vignette explains the coercion layer in detail.
* The `explorer-app` vignette walks through `hb_run_explorer()`.
